---
output:
  word_document: default
  html_document: default
---
#   Module 3 - Assignment 1
##  Jordan Roberts
### BAN 502

```{r, include=FALSE}
library(tidyverse)
library(MASS)
library(caret)

bike = read.csv("hour.csv")
```

```{r}
bike = bike %>% mutate(season = as_factor(as.character(season))) %>%
  mutate(season = fct_recode(season,
                             "Spring" = "1",
                             "Summer" = "2",
                             "Fall" = "3",
                             "Winter" = "4")) %>%
  mutate(yr = as_factor(as.character(yr)), mnth = as_factor(as.character(mnth)), hr = as_factor(as.character(hr))) %>%
  mutate(holiday = as_factor(as.character(holiday))) %>%
  mutate(holiday = fct_recode(holiday,
                              "NotHoliday" = "0",
                              "Holiday" = "1")) %>%
  mutate(workingday = as_factor(as.character(workingday))) %>%
  mutate(workingday = fct_recode(workingday,
                                 "NotWorkingDay" = "0",
                                 "WorkingDay" = "1")) %>%
  mutate(weathersit = as_factor(as.character(weathersit))) %>%
  mutate(weathersit = fct_recode(weathersit,
                                 "NoPrecip" = "1",
                                 "Misty" = "2",
                                 "LightPrecip" = "3",
                                 "HeavyPrecip" = "4")) %>%
  mutate(weekday = as_factor(as.character(weekday))) %>%
  mutate(weekday = fct_recode(weekday,
                              "Sunday" = "0",
                              "Monday" = "1",
                              "Tuesday" = "2",
                              "Wednesday" = "3",
                              "Thursday" = "4",
                              "Friday" = "5",
                              "Saturday" = "6"))
```

```{r}
train.rows = createDataPartition(y=bike$count, p=0.7, list=FALSE)
train = bike[train.rows,]
test = bike[-train.rows,]
set.seed(1234)
```

The training set has 12168 rows (~70% of 17379) while the test set has 5211 (~30%).  

```{r}
train.mod = lm(count ~ season+mnth+hr+holiday+weekday+temp+weathersit, train)
summary(train.mod)
```

The training model has an R-squared value of .62, fairly high and indicative of a good possibility of some linear fit.  

```{r}
train.predict = predict(train.mod)
head(train.predict)
```

With an intercept of -82.2, we'd expect values around the ones seen here. Note that 1 is missing from this set; we would expect to find 1 in the 30% test set that we'll find next.  

```{r}
test.predict = predict(train.mod, newdata=test)
head(test.predict)
```

Just like with the training predictions, these predictions are close to what we would expect. We found prediction 1 here, as well.  

```{r}
SSE = sum((test$count - test.predict)^2)
SST = sum((test$count - mean(test$count))^2)
1 - SSE/SST
```

Our original training R-squared value was 0.62, and our test sample returned a very close value. This indicates that our original model was a good fit.  

With this model, we've split the data at random into two groups, one containing 70% and the other containing 30% of the bike data frame. If we were to use k-fold cross-validation, we would be splitting the data into several groups or "folds" and testing combinations of them. For example, with 5 folds, we would have 4 folds make up the training set while the 5th is the test set. Then, we would do it again with a different fold as the test set and the remainders as the new training set. This would continue until we have analyzed each combination. This gives us a more thorough analysis of model performance.