---
output:
  word_document: default
  html_document: default
---
#   Module 2 - Assignment 2
##  Jordan Roberts
### BAN 502

```{r, include=FALSE}
library(tidyverse)
library(GGally)
library(MASS)
library(car)
library(gridExtra)
```

```{r}
bike = read.csv("hour.csv")

bike = bike %>% mutate(season = as_factor(as.character(season))) %>%
  mutate(season = fct_recode(season,
                             "Spring" = "1",
                             "Summer" = "2",
                             "Fall" = "3",
                             "Winter" = "4")) %>%
  mutate(yr = as_factor(as.character(yr)), mnth = as_factor(as.character(mnth)), hr = as_factor(as.character(hr))) %>%
  mutate(holiday = as_factor(as.character(holiday))) %>%
  mutate(holiday = fct_recode(holiday,
                              "NotHoliday" = "0",
                              "Holiday" = "1")) %>%
  mutate(workingday = as_factor(as.character(workingday))) %>%
  mutate(workingday = fct_recode(workingday,
                                 "NotWorkingDay" = "0",
                                 "WorkingDay" = "1")) %>%
  mutate(weathersit = as_factor(as.character(weathersit))) %>%
  mutate(weathersit = fct_recode(weathersit,
                                 "NoPrecip" = "1",
                                 "Misty" = "2",
                                 "LightPrecip" = "3",
                                 "HeavyPrecip" = "4")) %>%
  mutate(weekday = as_factor(as.character(weekday))) %>%
  mutate(weekday = fct_recode(weekday,
                              "Sunday" = "0",
                              "Monday" = "1",
                              "Tuesday" = "2",
                              "Wednesday" = "3",
                              "Thursday" = "4",
                              "Friday" = "5",
                              "Saturday" = "6"))
```

```{r}
ggcorr(bike, label=TRUE)
```

Temp and ATemp each have a correlation of 0.4, the highest for quantitative variables in the data frame.  
```{r}
grid.arrange(
  ggplot(bike, aes(season, count)) +
    geom_boxplot(),
  ggplot(bike, aes(holiday, count)) +
    geom_boxplot(),
  ggplot(bike, aes(weekday, count)) +
    geom_boxplot(),
  ggplot(bike, aes(workingday, count)) +
    geom_boxplot(),
  ggplot(bike, aes(weathersit, count)) +
    geom_boxplot(),  
  ggplot(bike, aes(yr, count)) +
    geom_boxplot(),
  ggplot(bike, aes(mnth, count)) +
    geom_boxplot(),
  ggplot(bike, aes(hr, count)) +
    geom_boxplot())
```

season - biking seems to pick up in the Summer and Fall seasons, then decrease toward Winter and Spring. This is likely due to the fairer weather.  

mnth - this seems to follow the same pattern as season, as months would be correlated with seasons in record. biking picks up toward April/May and continues strong until October/November, when weather gets worse.  

holiday - there isn't much change between the two options, though on holidays the average is a bit lower, likely due to staying home.  

weekday - not much change between the days besides the weekends, when count is lower than the rest of the week. likely also due to people staying home from work.  

workingday - similarly, non-working days see lower counts as well.  

weathersit - as expected, heavier precipitation correlates with substantially fewer bikers.  

yr - count increased in the second year, likely due to growth of the service and new customers adding to the existing ones.  

hr - biking picks up around 7-8 am, then again around 5-6 pm. this fits nicely into the typical work day, when commuters head out or head home.  
```{r}
bike2 = bike %>% dplyr::select(-c(instant, dteday, registered, casual))

emptymod = lm(count ~1, bike2)
fullmod = lm(count ~., bike2)

forwardmod = stepAIC(emptymod, direction = "forward", scope=list(upper=fullmod, lower=emptymod),trace=TRUE)
```

```{r}
summary(forwardmod)
```

The forward model includes hr, atemp, yr, weathersit, season, mnth, hum, weekday, holiday, windspeed, and temp. Many variables have negative coefficients, suggesting some multicollinearity. However, some variables, like weather, may be correct in their coefficients, as heavier precipitation has shown to decrease bike count. Most of the variables are significant.  
```{r}
backwardmod = stepAIC(fullmod, direction = "backward", trace = TRUE)
summary(backwardmod)
```

The backward model is the same as the forward, only missing count (obviously) and workingday. workingday is correlated with weekday, with Monday through Friday being typical workdays in the US. This would make the inclusion of workingday redundant.  

The model is quite large, but it shows what we'd gleaned from the plots in the beginning of this exercise. The code takes much less time to set up, however, so it could be more useful. It seems a strong indicator of correlation for this dataset. Many of the predictors appear to be correlated with each other, though, such as season and month. Care must be taken to properly evaluate the results of these models.